---
title: "Dynamic Exposure"
output: html_document
---

```{r}
knitr::opts_chunk$set(warning = FALSE)
```

# Data loading


```{r}
library(data.table)

# Reading the data and identifying individuals and their first shooting event
njforce <- data.table::fread("../data-raw/njforce_200210.csv")
reports <- subset(
  njforce,
  select = c(
    date, officerid, firearm_discharged, firearm_pointed, incidentid,
    officer_male, officer_nyears, officer_race, officer_rank, town,
    officer_po, officer_sleo, nsubjects, Incident_type
    ))
reports[, date := as.Date(date, format = "%m/%d/%Y")]
```

I don't see many discharged reports, so we will look at pointed. Only 11 records
show discharged weapon, though 764 are firearm pointed. For the 11 records, it
would suffice do a qualitative analysis.

```{r}
table(reports$firearm_pointed, reports$firearm_discharged)
```

How many officers per incident?

```{r}
# How many officers per incident
tab <- table(njforce[, list(N=.N), by = incidentid]$N)
tab <- cbind(
  N = tab,
  pcent = tab/sum(tab),
  cum_pcent = cumsum(tab/sum(tab))
  )
knitr::kable(tab)
```


# Calculating Dynamic Exposure

```{r exposure-preamble, cache = TRUE}
# Counting the number of guns pointed
reports[, nofficers := .N, by = "incidentid"]

# Ordering the data for some cumulative computations
reports <- reports[order(officerid, date),]

# Event number counter
reports[, nevent := 1:.N, by = "officerid"]

# Creating a cumulative counter per officer
reports[, fdischarged := cumsum(firearm_discharged), by = "officerid"]
reports[, fpointed    := cumsum(firearm_pointed), by = "officerid"]
reports[, fpointed    := cumsum(firearm_pointed), by = "officerid"]

# TRUE if the individual had ever pointed a gun in the past
reports[, ever_pointed := fpointed != 0]
```

**Immediate Exposure**: Number of officers other than $i$ who pointed a gun at event $T$.

\begin{equation}
E_{iT}^I = \sum_{j \neq i}x_{iT}
\end{equation}

```{r exposure-i}
reports[, exposure_i := sum(firearm_pointed) - firearm_pointed, by = "incidentid"]
reports[, exposure_i := shift(exposure_i, type="lag"), by = "officerid"]
```

**Deferred Exposure**: Number of officers other than $i$ who participated in event $T$ who have ever pointed a gun.

\begin{equation}
E_{iT}^D = \sum_{j\neq i}\max_{t\leq T}x_{jt}
\end{equation}

```{r exposure-d}
reports[, exposure_d := sum(ever_pointed) - ever_pointed, by = "incidentid"]
reports[, exposure_d := shift(exposure_d, type="lag"), by = "officerid"]
```

**Cumulative Exposure**: Number of officers other than $i$ who participated in event $T$ who have ever pointed a gun in the presence of $i$. 

\begin{equation}
E_{iT}^C = \sum_{j \neq i}\max_{t\leq T, t\in T(i)}x_{iT}
\end{equation}

Where $T(i)$ is the set that represent the events in which $i$ was involved.

```{r exposure-c}
#TBD
```


```{r exposures}
# Identifying the first event with either pointed or discharged, we will
# use this later.
reports[, first := ever_pointed & !shift(ever_pointed, type="lag", fill=TRUE), by = "officerid"]

# Average peer experience (as in number of reported events)
reports[, peer_exp := (sum(nevent) - nevent)/(nofficers - 1 + 1e-20), by = "incidentid"]

# Adding other features
reports[, female := as.integer(officer_male == 0)]
reports[, police := as.integer(officer_po == 1)]
reports[, sleo   := as.integer(officer_sleo == 1)]

# Indicator variable to flag whether it is relevant or not
reports[, relevant := (1:.N > 1L) & (!ever_pointed | first), by = "officerid"]

# Violent or weapon related incident
reports[, violent_or_weapon := as.integer(grepl("armed|gun|violen|weap|shoot|knife", Incident_type))]
```

# Naive models

## Predicting the first incident

```{r results='asis'}
reports[, exposure_yesno := as.integer(exposure_d > 0)]
reports[, pressure := nofficers/(nsubjects + 1e-20)]

ans0 <- glm(
  firearm_pointed ~ exposure_yesno + nevent + nofficers + female +
    officer_nyears + factor(officer_race) + violent_or_weapon + factor(town),
  data   = reports,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans1 <- glm(
  firearm_pointed ~ exposure_yesno + I(exposure_yesno * female) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + violent_or_weapon +
    factor(town),
  data   = reports,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans2 <- glm(
  firearm_pointed ~ exposure_yesno + I(exposure_yesno * female) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + violent_or_weapon + factor(town),
  data   = reports,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans3 <- glm(
  firearm_pointed ~ exposure_d + I(exposure_d * female) + I(exposure_d^2) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + violent_or_weapon + factor(town),
  data   = reports,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans4 <- glm(
  firearm_pointed ~ exposure_d +  I(exposure_d^2) + 
    sleo + nsubjects + violent_or_weapon + factor(town),
  data   = reports,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans5 <- glm(
  firearm_pointed ~ exposure_d +  I(exposure_d^2) + 
    sleo + nsubjects + factor(town),
  data   = reports,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans6 <- glm(
  firearm_pointed ~ exposure_d +  I(exposure_d^2) + 
    sleo + nsubjects + factor(town) + pressure,
  data   = reports,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

# reports$officer_race

cmap <- list(
  exposure_yesno           = "Exposure (yes/no)",
  `I(exposure_yesno * female)` = "Exposure (yes/no) x Female",
  exposure_d               = "Exposure (count)",
  `I(exposure_d^2)`        = "Exposure (count) ^2",
  `I(exposure_d * female)` = "Exposure (count) x Female",
  exposure_d_indirect      = "Exposure (indirect, count)",
  nevent                   = "Num. of Event",
  nofficers                  = "N. Officers Involved",
  officer_nyears           = "Tenure (t)",
  nsubjects                = "N. Subjects",
  peer_exp                 = "Peer experience (avg. years)",
  female                   = "Female (yes/no)",
  `factor(officer_race)1`  = "Asian (yes/no)", 
  `factor(officer_race)2`  = "Black (yes/no)", 
  `factor(officer_race)3`  = "Hispanic (yes/no)", 
  `factor(officer_race)4`  = "White (yes/no)", 
  police                   = "is police off. (yes/no)",
  sleo                     = "is law enforcement off. (yes/no)",
  violent_or_weapon        = "Violent or Weapon related (yes/no)",
  pressure                 = "N Officers/N Subjects"
)

# Officer_race
# 1 = asian
# 2 = black
# 3 = Hispanic
# 4 = white

library(texreg)
models <- mget(ls(pattern = "^ans[0-9]+$"))

# Computing McFadden's pseudo r2
pr2 <- sapply(models, function(m) {
  1 - with(m, deviance/null.deviance)
})

auc <- sapply(models, function(m) {
  AUC::auc(AUC::roc(predict(m, type = "response"), as.factor(m$y)))
})

htmlreg(
  models, custom.coef.map = cmap,
  custom.gof.rows = list("Pseudo R2" = pr2, "AUC" = auc)
  )
# screenreg(models, custom.coef.map = cmap,custom.gof.rows = list("Pseudo R2" = pr2))
```


## Predicting any incident

```{r results='asis'}
library(lme4)

# Lagging self-exposure
reports[, firearm_pointed_prev := shift(firearm_pointed, type = "lag", fill = TRUE), by = "officerid"]
reports[, firearm_pointed_prev := cumsum(firearm_pointed_prev), by = "officerid"]

# reports[, nevents_tot := .N, by = "officerid"]
reports[, relevant_any := 1:.N > 1, by = "officerid"]

reports[, exposure_yesno := as.integer(exposure_d > 0)]
```

```{r fit-ans0, cache=TRUE}
ans0 <- glmer(
  firearm_pointed ~ exposure_yesno + nevent + nofficers + female +
    officer_nyears + factor(officer_race) + firearm_pointed_prev + violent_or_weapon +
    factor(town) + (1 | officerid),
  data   = reports,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )
```


```{r fit-ans1, cache=TRUE}
ans1 <- glmer(
  firearm_pointed ~ exposure_yesno + I(exposure_yesno * female) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + firearm_pointed_prev +
    violent_or_weapon + factor(town) + (1 | officerid),
  data   = reports,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )
```

```{r fit-ans2, cache=TRUE}
ans2 <- glmer(
  firearm_pointed ~ exposure_yesno + I(exposure_yesno * female) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + firearm_pointed_prev + violent_or_weapon +
    factor(town) + (1 | officerid),
  data   = reports,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )
```

```{r fit-ans3, cache=TRUE}
ans3 <- glmer(
  firearm_pointed ~ exposure_d + I(exposure_d * female) + I(exposure_d^2) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + firearm_pointed_prev + violent_or_weapon +
    factor(town) + (1 | officerid),
  data   = reports,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )
```

```{r fit-ans4, cache=TRUE}
ans4 <- glmer(
  firearm_pointed ~ exposure_d + I(exposure_d * female) + I(exposure_d^2) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + firearm_pointed_prev + violent_or_weapon +
    factor(town) + (1 | officerid),
  data   = reports,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )
```

```{r fit-ans5, cache=TRUE}
ans5 <- glmer(
  firearm_pointed ~ exposure_d + I(exposure_d * female) + I(exposure_d^2) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + firearm_pointed_prev + violent_or_weapon +
    factor(town) + peer_exp + (1 | officerid),
  data   = reports,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )
```

```{r fit-ans6, cache=TRUE}
ans6 <- glmer(
  firearm_pointed ~ exposure_d + I(exposure_d^2) + nevent +
    officer_nyears + 
    sleo + nsubjects + firearm_pointed_prev + violent_or_weapon +
    factor(town) + (1 | officerid),
  data   = reports,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )
```

```{r fit-ans7, cache=TRUE}
ans7 <- glmer(
  firearm_pointed ~ exposure_d + I(exposure_d^2) + nevent +
    officer_nyears + 
    sleo + nsubjects + firearm_pointed_prev + factor(town) + (1 | officerid),
  data   = reports,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )
```

```{r fit-ans8, cache=TRUE}
ans8 <- glmer(
  firearm_pointed ~ exposure_d + I(exposure_d^2) + nevent +
    officer_nyears + pressure +
    sleo + nsubjects + firearm_pointed_prev + factor(town) + (1 | officerid),
  data   = reports,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )
```

```{r results='asis'}
cmap <- c(cmap, list(firearm_pointed_prev = "N Previous Pointed"))

# Officer_race
# 1 = asian
# 2 = black
# 3 = Hispanic
# 4 = white

library(texreg)
models <- mget(ls(pattern = "^ans[0-9]+$"))

# Computing McFadden's pseudo r2
# pr2 <- sapply(models, function(m) {
#   1 - with(m, deviance/null.deviance)
# })

auc <- sapply(models, function(m) {
  AUC::auc(AUC::roc(predict(m, type = "response"), as.factor(m@resp$y)))#as.factor(m$y)))
})

htmlreg(
  models, custom.coef.map = cmap,
  custom.gof.rows = list(#"Pseudo R2" = pr2,
    "AUC" = auc)
  )
# screenreg(models, custom.coef.map = cmap,custom.gof.rows = list("Pseudo R2" = pr2))
```


Throw in time of the day.

Time of the year.

See if we can get info about the area.

Broker in misconduct network.
