---
title: "Dynamic Exposure"
output: html_document
---

```{r}
knitr::opts_chunk$set(warning = FALSE)
```

# Overview

Exposure dynamically until the first firearm-related incident. In particular, we
use the following definition:

Given that there was an incident, what is the effect of exposure to previous firearm
incidents (indirect exposure)?

Question: What type of events are the most relevant for the analysis? Is there a
way to classify the events?


```{r}
library(data.table)

# Reading the data and identifying individuals and their first shooting event
njforce <- data.table::fread("../data-raw/njforce_200210.csv")
incidents <- subset(
  njforce,
  select = c(
    date, officerid, firearm_discharged, firearm_pointed, incidentid,
    officer_male, officer_nyears, officer_race, officer_rank, town,
    officer_po, officer_sleo, nsubjects, Incident_type
    ))
incidents[, date := as.Date(date, format = "%m/%d/%Y")]
```

I don't see many discharged incidents, so we will look at pointed. Only 11 records
show discharged weapon, though 764 are firearm pointed. For the 11 records, it
would suffice do a qualitative analysis.

```{r}
table(incidents$firearm_pointed, incidents$firearm_discharged)
```

How many officers per incident?

```{r}
# How many officers per incident
tab <- table(njforce[, list(N=.N), by = incidentid]$N)
tab <- cbind(
  N = tab,
  pcent = tab/sum(tab),
  cum_pcent = cumsum(tab/sum(tab))
  )
knitr::kable(tab)
```


# Calculating Dynamic Exposure



1.  Compute the number of events until the date (running count), this will also
    imply tagging officers since when they were firearm-involved.

2.  Then, for each individual+event, we compute exposure to that date:
    
    a. Number of its neighbors that were exposed to a firearm related event.
    
    b. a

```{r exposure-calc, cache = TRUE}
# Counting the number of guns pointed
incidents[, nofficers := .N, by = "incidentid"]

# Ordering the data for some cumulative computations
incidents <- incidents[order(officerid, date),]

# Event number counter
incidents[, nevent := 1:.N, by = "officerid"]

# Creating a cumulative counter per officer
incidents[, fdischarged := cumsum(firearm_discharged), by = "officerid"]
incidents[, fpointed    := cumsum(firearm_pointed), by = "officerid"]
incidents[, fpointed    := cumsum(firearm_pointed), by = "officerid"]
incidents[, exposed     := fpointed != 0]

# Identifying the first event with either pointed or discharged, we will
# use this later.
incidents[, first := exposed & !shift(exposed, type="lag", fill=TRUE), by = "officerid"]

# Average peer experience (as in number of reported events)
incidents[, peer_exp := (sum(nevent) - nevent)/(nofficers - 1 + 1e-20), by = "incidentid"]

# Iterating through each entry
incidents$exposure_t <- 0L
incidents$exposure_t_indirect <- 0L
# i<-5
for (i in 1:nrow(incidents)) {
  
  # Getting the officer id and the date
  officer_i <- incidents$officerid[i]
  date_i    <- incidents$date[i]
  
  # Subset the data to all incidents in which officer i was involved up to
  # the i-th date
  incidents_i <- incidents[officerid == officer_i & date < date_i]$incidentid
  incidents_i <- incidents[(incidentid %in% incidents_i) & officerid != officer_i]

  if (nrow(incidents_i) > 0L) {
    indirect_exposure <- incidents_i[, list(count = max(exposed)), by="officerid"]
    direct_exposure   <- incidents_i[, list(count = max(firearm_pointed)), by="officerid"]
    incidents$exposure_t_indirect[i] <- sum(indirect_exposure$count)
    incidents$exposure_t[i]          <- sum(direct_exposure$count)
  }
  
  if (!(i %% 100))
    message(i, " out of ", nrow(incidents), " done.")
  
}

# Adding other features
incidents[, female := as.integer(officer_male == 0)]
incidents[, police := as.integer(officer_po == 1)]
incidents[, sleo   := as.integer(officer_sleo == 1)]

# Indicator variable to flag whether it is relevant or not
incidents[, relevant := (1:.N > 1L) & (!exposed | first), by = "officerid"]

# Violent or weapon related incident
incidents[, violent_or_weapon := as.integer(grepl("armed|gun|violen|weap|shoot|knife", Incident_type))]
```

# Naive models

## Predicting the first incident

```{r results='asis'}
incidents[, exposure_yesno := as.integer(exposure_t > 0)]
incidents[, pressure := nofficers/(nsubjects + 1e-20)]

ans0 <- glm(
  firearm_pointed ~ exposure_yesno + nevent + nofficers + female +
    officer_nyears + factor(officer_race) + violent_or_weapon + factor(town),
  data   = incidents,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans1 <- glm(
  firearm_pointed ~ exposure_yesno + I(exposure_yesno * female) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + violent_or_weapon +
    factor(town),
  data   = incidents,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans2 <- glm(
  firearm_pointed ~ exposure_yesno + I(exposure_yesno * female) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + violent_or_weapon + factor(town),
  data   = incidents,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans3 <- glm(
  firearm_pointed ~ exposure_t + I(exposure_t * female) + I(exposure_t^2) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + violent_or_weapon + factor(town),
  data   = incidents,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans4 <- glm(
  firearm_pointed ~ exposure_t +  I(exposure_t^2) + 
    sleo + nsubjects + violent_or_weapon + factor(town),
  data   = incidents,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans5 <- glm(
  firearm_pointed ~ exposure_t +  I(exposure_t^2) + 
    sleo + nsubjects + factor(town),
  data   = incidents,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans6 <- glm(
  firearm_pointed ~ exposure_t +  I(exposure_t^2) + 
    sleo + nsubjects + factor(town) + pressure,
  data   = incidents,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

ans7 <- glm(
  firearm_pointed ~ exposure_t_indirect +  I(exposure_t_indirect^2) + 
    sleo + nsubjects + factor(town) + pressure,
  data   = incidents,
  subset = relevant == TRUE,
  family = binomial(link="logit")
  )

# incidents$officer_race

cmap <- list(
  exposure_yesno           = "Exposure (yes/no)",
  `I(exposure_yesno * female)` = "Exposure (yes/no) x Female",
  exposure_t               = "Exposure (count)",
  `I(exposure_t^2)`        = "Exposure (count) ^2",
  `I(exposure_t * female)` = "Exposure (count) x Female",
  exposure_t_indirect      = "Exposure (indirect, count)",
  `I(exposure_t_indirect^2)` = "Exposure (indirect, count) ^2",
  nevent                   = "Num. of Event",
  nofficers                  = "N. Officers Involved",
  officer_nyears           = "Tenure (t)",
  nsubjects                = "N. Subjects",
  peer_exp                 = "Peer experience (avg. years)",
  female                   = "Female (yes/no)",
  `factor(officer_race)1`  = "Asian (yes/no)", 
  `factor(officer_race)2`  = "Black (yes/no)", 
  `factor(officer_race)3`  = "Hispanic (yes/no)", 
  `factor(officer_race)4`  = "White (yes/no)", 
  police                   = "is police off. (yes/no)",
  sleo                     = "is law enforcement off. (yes/no)",
  violent_or_weapon        = "Violent or Weapon related (yes/no)",
  pressure                 = "N Officers/N Subjects"
)

# Officer_race
# 1 = asian
# 2 = black
# 3 = Hispanic
# 4 = white

library(texreg)
models <- mget(ls(pattern = "^ans[0-9]+$"))

# Computing McFadden's pseudo r2
pr2 <- sapply(models, function(m) {
  1 - with(m, deviance/null.deviance)
})

auc <- sapply(models, function(m) {
  AUC::auc(AUC::roc(predict(m, type = "response"), as.factor(m$y)))
})

htmlreg(
  models, custom.coef.map = cmap,
  custom.gof.rows = list("Pseudo R2" = pr2, "AUC" = auc)
  )
# screenreg(models, custom.coef.map = cmap,custom.gof.rows = list("Pseudo R2" = pr2))
```


## Predicting any incident

```{r results='asis'}

# Lagging self-exposure
incidents[, firearm_pointed_prev := shift(firearm_pointed, type = "lag", fill = TRUE), by = "officerid"]
incidents[, firearm_pointed_prev := cumsum(firearm_pointed_prev), by = "officerid"]

incidents[, relevant_any := 1:.N > 1, by = "officerid"]

incidents[, exposure_yesno := as.integer(exposure_t > 0)]
ans0 <- glm(
  firearm_pointed ~ exposure_yesno + nevent + nofficers + female +
    officer_nyears + factor(officer_race) + firearm_pointed_prev + violent_or_weapon +
    factor(town),
  data   = incidents,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )

ans1 <- glm(
  firearm_pointed ~ exposure_yesno + I(exposure_yesno * female) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + firearm_pointed_prev +
    violent_or_weapon + factor(town),
  data   = incidents,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )

ans2 <- glm(
  firearm_pointed ~ exposure_yesno + I(exposure_yesno * female) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + firearm_pointed_prev + violent_or_weapon +
    factor(town),
  data   = incidents,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )

ans3 <- glm(
  firearm_pointed ~ exposure_t + I(exposure_t * female) + I(exposure_t^2) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + firearm_pointed_prev + violent_or_weapon +
    factor(town),
  data   = incidents,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )

ans4 <- glm(
  firearm_pointed ~ exposure_t + I(exposure_t * female) + I(exposure_t^2) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + firearm_pointed_prev + violent_or_weapon +
    factor(town),
  data   = incidents,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )

ans5 <- glm(
  firearm_pointed ~ exposure_t + I(exposure_t * female) + I(exposure_t^2) + nevent +
    nofficers + female + officer_nyears + factor(officer_race) + 
    police + sleo + nsubjects + firearm_pointed_prev + violent_or_weapon +
    factor(town) + peer_exp,
  data   = incidents,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )

ans6 <- glm(
  firearm_pointed ~ exposure_t + I(exposure_t^2) + nevent +
    officer_nyears + 
    sleo + nsubjects + firearm_pointed_prev + violent_or_weapon +
    factor(town),
  data   = incidents,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )

ans7 <- glm(
  firearm_pointed ~ exposure_t + I(exposure_t^2) + nevent +
    officer_nyears + 
    sleo + nsubjects + firearm_pointed_prev + factor(town),
  data   = incidents,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )

ans8 <- glm(
  firearm_pointed ~ exposure_t + I(exposure_t^2) + nevent +
    officer_nyears + pressure +
    sleo + nsubjects + firearm_pointed_prev + factor(town),
  data   = incidents,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )

ans9 <- glm(
  firearm_pointed ~ exposure_t_indirect + I(exposure_t_indirect^2) + nevent +
    officer_nyears + pressure +
    sleo + nsubjects + firearm_pointed_prev + factor(town),
  data   = incidents,
  subset = relevant_any == TRUE,
  family = binomial(link="logit")
  )

# incidents$officer_race

cmap <- c(cmap, list(firearm_pointed_prev = "N Previous Pointed"))

# Officer_race
# 1 = asian
# 2 = black
# 3 = Hispanic
# 4 = white

library(texreg)
models <- mget(ls(pattern = "^ans[0-9]+$"))

# Computing McFadden's pseudo r2
pr2 <- sapply(models, function(m) {
  1 - with(m, deviance/null.deviance)
})

auc <- sapply(models, function(m) {
  AUC::auc(AUC::roc(predict(m, type = "response"), as.factor(m$y)))
})

htmlreg(
  models, custom.coef.map = cmap,
  custom.gof.rows = list("Pseudo R2" = pr2, "AUC" = auc)
  )
# screenreg(models, custom.coef.map = cmap,custom.gof.rows = list("Pseudo R2" = pr2))
```


Throw in time of the day.

Time of the year.

See if we can get info about the area.

Broker in misconduct network.
