---
title: "Model 1"
output: html_document
---

```{r parameters}
max_n_events <- 15L
```


```{r load-data}
library(data.table)

# Reading the data and identifying individuals and their first shooting event
njforce <- data.table::fread("data-raw/njforce_200820_clean.csv")
reports <- subset(
  njforce,
  select = c(
    date, officerid_num, firearm_discharged, firearm_pointed, incidentid,
    officer_male, officer_nyears, officer_race, officer_rank, town,
    officer_po, officer_sleo, nsubjects, Incident_type
    ))
reports[, date := as.Date(date, format = "%m/%d/%Y")]
colnames(reports)[2] <- "officerid"
```


```{r exposure-i}
reports <- reports[order(officerid, incidentid),]
reports[, exposure_i := as.integer((sum(firearm_pointed) - firearm_pointed) > 0), by = "incidentid"]
reports[, exposure_i := shift(exposure_i, type="lag"), by = "officerid"]
```

```{r more-features}
# Adding other features
reports[, female := as.integer(officer_male == 0)]
reports[, police := as.integer(officer_po == 1)]
reports[, sleo   := as.integer(officer_sleo == 1)]

# Violent or weapon related incident
reports[, violent_or_weapon := as.integer(grepl("armed|gun|violen|weap|shoot|knife", Incident_type))]

# N officers 
reports[, nofficers := .N, by = incidentid]

# Count of incidents involving more than one (cumulative)
reports[, prev_no_solo := nofficers > 1, by = officerid]
reports[, prev_no_solo := shift(prev_no_solo, type = "lag", fill = FALSE), by = officerid]

# Number of events (ever)
reports[, nevent := 1:.N, by = officerid]
reports[, nofficers := .N, by = incidentid]
reports[, nevents := .N, by = officerid]

# Previous event
reports[, firearm_pointed_prev := shift(firearm_pointed, type = "lag", fill = FALSE), by = officerid]

if (FALSE) {
  View(
    subset(reports, select = c(
      officerid, incidentid, nofficers, prev_no_solo, date
      )))
}

reports[, townid := as.integer(as.factor(town))]
reports[, raceid := as.integer(as.factor(officer_race))]
```


# Baseline model

Likelihood of pointing a firearm given the previous incident they participated they were not alone. For the model, we will only look at individuals who potentially were exposed to a firearm incident in which other individuals participated.

```{r dist-prefilter}
with(
  reports[prev_no_solo == TRUE & nevents <= 10],
  table(firearm_pointed, exposure_i)
)
table(reports[, unique(nevents), by = officerid]$V1)
```


```{r baseline}
ans0 <- glm(
  firearm_pointed ~ 
    exposure_i + female + police + officer_nyears + factor(town) +
    nevent + factor(officer_race) + nofficers + firearm_pointed_prev,
  data = reports,
  subset = prev_no_solo & nevents <= max_n_events
)

ans1 <- glm(
  firearm_pointed ~ 
    exposure_i + female + police + officer_nyears + factor(town) +
    nevent+ factor(officer_race) + violent_or_weapon + nofficers + firearm_pointed_prev,
  data = reports,
  subset = prev_no_solo & nevents <= max_n_events
)

ans2 <- glm(
  firearm_pointed ~ 
    exposure_i*female + police + officer_nyears + factor(town) +
    nevent+ factor(officer_race) + violent_or_weapon + nofficers + firearm_pointed_prev,
  data = reports,
  subset = prev_no_solo & nevents <= max_n_events
)
```



There are a couple of dozen officers who will show

```{r re-model, cache=FALSE}
library(lme4)
ans_re <- glmer(
  firearm_pointed ~ 
    exposure_i + female + police + officer_nyears + factor(town) +
    nevent+ factor(officer_race) + (1 | officerid) + nofficers + firearm_pointed_prev,
  data   = reports,
  subset = prev_no_solo & nevents <= max_n_events,
  family = binomial(link="logit")
  )
```


```{r baseline-table}
texreg::screenreg(list(ans0, ans1, ans2, ans_re), stars = c(.001, .01, .05, .1))
```

# Permutation test

For the permutation test, one strategy was propose swaps based on the officer id (so within the officer) and whether the event included multiple officers or not. Unfortunately (or thankfully), there are many records for which we were not able to find a suitable match. Because of that reason, we opted for taking a different approach, and this is, find matches looking at the following variables: gender, town, whether the event was with multiple officers or not.

```{r candidates}
library(njforce)
# reports <- reports[(nofficers == 1) & prev_no_solo,]
reports[, datenum := as.integer(date)]
reports[, mult_officers := as.integer(nofficers > 1)]
x <- as.matrix(
  subset(reports, select = c(female, townid, mult_officers, prev_no_solo, datenum, raceid))
)

candidates <- find_candidates(
  features = x,
  upper    = c(0, 0, 0, 0, 45, 0),
  lower    = c(0, 0, 0, 0, -45, 0),
  as_abs   = c(TRUE, TRUE, TRUE, TRUE, FALSE, TRUE)
)

library(magrittr)
reports[, which(!prev_no_solo & nevents < max_n_events)] %>%
  extract(candidates, .) %>%
  sapply(length) %>%
  table
```

```{r perm-test1}
Nperm <- 5000

# Precomputing permutations
set.seed(144)
permutations <- replicate(
  Nperm, permute(candidates),
  simplify = FALSE)

# How many will change
permutations_m <- do.call(cbind, permutations)[reports$prev_no_solo & reports$nevents <= max_n_events,]
permutations_m <- colSums(permutations_m != permutations_m[,1])/nrow(permutations_m)
hist(
  permutations_m[-1],
  main = "Proportion of permuted rows",
  breaks = 50
  )
```

```{r perm-test1-fit}
# Baseline model
baseline_model1 <- firearm_pointed ~ exposure_i * female + police +
  officer_nyears + factor(town) + nevent + factor(officer_race) + nofficers  + firearm_pointed_prev

ans_baseline1 <- glm(baseline_model1,
  data = reports,
  subset = prev_no_solo & nevents <= max_n_events
)

# Preparing the cluster
cl <- parallel::makePSOCKcluster(4)
parallel::clusterExport(cl, c("reports", "baseline_model1", "max_n_events"))
invisible(parallel::clusterEvalQ(cl, library(data.table)))

# Running the analysis
ans_perm1 <- parallel::parLapply(cl, permutations, function(p) {
  
  # Reorder the data
  tmp <- data.table::copy(reports)
  tmp$firearm_pointed <- tmp$firearm_pointed[p + 1]
  
  # environment(baseline_model) <- environment()
  tmp <- glm(
    baseline_model1, data = tmp,
    subset = prev_no_solo & nevents <= max_n_events)
  
  coef(tmp)
})

# Baseline model
baseline_model2 <- firearm_pointed ~ exposure_i + police + female +
  officer_nyears + factor(town) + nevent + factor(officer_race) + nofficers  + firearm_pointed_prev

ans_baseline2 <- glm(baseline_model2,
  data = reports,
  subset = prev_no_solo & nevents <= max_n_events
)

parallel::clusterExport(cl, "baseline_model2")

ans_perm2 <- parallel::parLapply(cl, permutations, function(p) {
  
  # Reorder the data
  tmp <- data.table::copy(reports)
  tmp$firearm_pointed <- tmp$firearm_pointed[p + 1]
  
  # environment(baseline_model) <- environment()
  tmp <- glm(
    baseline_model2, data = tmp,
    subset = prev_no_solo & nevents <= max_n_events
    )
  
  coef(tmp)
})

# Closing the other sessions
parallel::stopCluster(cl)
```


```{r report, fig.width=7, fig.height=10}
ans_perm1 <- do.call(rbind, ans_perm1)
pvals <- rowMeans(t(ans_perm1) < coef(ans_baseline1))
pvals[] <- ifelse(pvals > .5, 1 - pvals, pvals) * 2
ans_sd <- apply(ans_perm1, 2, sd)

knitr::kable(cbind(Beta = coef(ans_baseline1), pval = pvals, sd = ans_sd), digits = 4)

ans_perm2 <- do.call(rbind, ans_perm2)
pvals <- rowMeans(t(ans_perm2) < coef(ans_baseline2))
pvals[] <- ifelse(pvals > .5, 1 - pvals, pvals) * 2
ans_sd <- apply(ans_perm2, 2, sd)

knitr::kable(cbind(Beta = coef(ans_baseline2), pval = pvals, sd = ans_sd), digits = 4)
```

```{r dist-plot, fig.width=6, fig.height=9}
op <- par(mar = par("mar") * c(1, 3, 1, 1))
cols <- c("gray", "steelblue" )[(pvals <= .05) + 1]
boxplot(
  t(t(ans_perm1)), horizontal = TRUE, las = 1,
  col    = cols,
  border = cols,
  main   = "Null distributions vs\nEstimated coefficients"
)
points(
  x = coef(ans_baseline1), y = 1:length(coef(ans_baseline1)), pch = 4, lwd=2, col="tomato"
  )
abline(v=0)
par(op)
```


```{r dist-plot2, fig.width=6, fig.height=9}
op <- par(mar = par("mar") * c(1, 3, 1, 1))
cols <- c("gray", "steelblue" )[(pvals <= .05) + 1]
boxplot(
  t(t(ans_perm2)), horizontal = TRUE, las = 1,
  col    = cols,
  border = cols,
  main   = "Null distributions vs\nEstimated coefficients"
)
points(
  x = coef(ans_baseline2), y = 1:length(coef(ans_baseline2)), pch = 4, lwd=2, col="tomato"
  )
abline(v=0)
par(op)
```

